---
title: "Gathering Data for the SPI Literature Review Paper"
author: "Brian Stacy"
date: "8/16/2023"
output: html_document
---

This file will gather data, mostly from the World Bank's World Development Indicators combine them with SPI data and data from other indicators of statistical performance, and save the resulting data as CSVs for use in the file that produces the final tables and figures.

Note: By re-running this code, you will overwrite all data used to produce the tables and figures for the paper, and thus the results may not replicate the final results in the paper.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	fig.height=6,
	fig.width = 8,
	dev=c("png","tiff")
)
library(data.table)
library(tidyverse)
library(flextable)
library(here)
library(haven)
library(wbstats)
library(readxl)
library(zoo)
#devtools::install_github("worldbank/pipr")
library(pipr) #World Bank PIP package

dir <- here()
raw_dir <- paste0(dir, "/01_raw_data/")
output_dir <- paste(dir, '/03_output/', sep="/")
#weights (either unity (1) or population)
wgt <- 1
end_date=2022
```


```{r}
#| label: pipdata

#read in PIP data for Figure 1 from the paper.
#new pipr package already includes survey coverage and comparability measures
pip_df <- get_stats()

# get mean consumption/income per day as of latest year (mean)
welfare_df <- pip_df %>%
  filter(survey_coverage %in% c('national')) %>% #keep just the national surveys
  filter(reporting_level %in% c('national')) %>% #keep just the national survey
  rename(iso3c=country_code,
         mean_welfare=mean) %>%
  group_by(iso3c) %>%
  filter(year==max(year)) %>%
  #keep consumption survey for Poland, Haiti, Angola, PHL
  filter(!(iso3c=="POL" & welfare_type=='income'))    %>%
  filter(!(iso3c=="HTI" & welfare_type=='income'))    %>%
  filter(!(iso3c=="AGO" & welfare_type=='income'))    %>%
  filter(!(iso3c=="PHL" & welfare_type=='income'))    %>%
  select(iso3c, mean_welfare)  

#reshape the data
pip_df <- pip_df %>%
  filter(survey_coverage %in% c('national')) %>% #keep just the national surveys
  filter(reporting_level %in% c('national')) %>% #keep just the national surveys
  filter(between(year, 1982, 2022)) %>% #keep data from 1982 to 2022
  #aggregate totals by country
  rename(iso3c=country_code) %>%
  group_by(iso3c) %>%
  summarise(
    number_surveys=n(),
    avg_mean_welfare = mean(mean)
  ) #get counts of the number of surveys by country 

#add welfare
pip_df <- pip_df %>%
  left_join(welfare_df)

#save
write_excel_csv(pip_df, paste0(output_dir, "/Figure_1_data.csv"))

```


```{r spidataload, echo=TRUE}
#add in population
pop_df <- wbstats::wb_data(country="all",
             indicator='SP.POP.TOTL',
             start_date=2004,
             end_date=end_date) %>%
  mutate(date=as.numeric(date)) %>%
  mutate(population=SP.POP.TOTL) %>%
  select(country, date, population) 
#import data
spi_index_df <- read_csv(paste(raw_dir, 'SPI_index.csv', sep="/")) %>%
    left_join(pop_df) 


spi_df_final <- read_csv( file = paste(raw_dir, 'SPI_data.csv', sep="/")) %>%
  left_join(pop_df) 
#metadata 
metadata <- data.frame(
  series = c('SPI.INDEX', 'SPI.INDEX.PIL1', 'SPI.INDEX.PIL2', 'SPI.INDEX.PIL3', 'SPI.INDEX.PIL4', 'SPI.INDEX.PIL5'),
  indicator_name = c('SPI Overall Score', 'Pillar 1: Data Use', 'Pillar 2: Data Services', 'Pillar 3: Data Products', 'Pillar 4: Data Sources', 'Pillar 5: Data Infrastructure')
)
#get list of country info
country_list <- wbstats::wb_countries()

write_excel_csv(country_list, paste0(raw_dir, "country_metadata.csv"))

SPI_end_date <- spi_index_df %>%
  filter(date==end_date) %>%
  filter(!is.na(SPI.INDEX) & !is.na(weights)) %>%
  mutate(ISO_A3_EH=iso3c) 


SPI <- spi_index_df
#read in TopoJSON from World Bank
#countries <- geojsonio::geojson_read("WB_countries_Admin0_lowres.geojson",
#                                     what = "sp")
```


```{r}
span <- c(2004:end_date)

#produce an empty dataset with the right number of rows (country X year observations)
spi_df_empty <- bind_rows(replicate(length(span), wbstats::wbcountries(), simplify = FALSE), .id='date') %>%
  mutate(date=as.numeric(date)+span[1]-1) %>%
  filter(region!="Aggregates") # take out the aggregates (LAC, SAR, etc)

```


```{r correlates}


start_date=2010

correlates_df <- wb_data(country="all", 
              indicator=c('SI.POV.DDAY', 'SI.POV.GINI'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('SI.POV.DDAY', 'SI.POV.GINI'), .direction = 'downup') %>%
  filter(date==end_date)


ge_df <- wb_data(country="all", 
              indicator=c('GE.EST'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('GE.EST'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c, date, GE.EST)

gdp_df <- wb_data(country="all", 
              indicator=c('NY.GDP.PCAP.KD'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('NY.GDP.PCAP.KD'), .direction = 'downup') %>%
  filter(date==end_date) %>%  select(iso3c, date, NY.GDP.PCAP.KD)

hci_df <- wb_data(country="all", 
              indicator=c('HD.HCI.OVRL'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('HD.HCI.OVRL'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c, HD.HCI.OVRL)

undernourish_df <- wb_data(country="all", 
              indicator=c('SN.ITK.DEFC.ZS'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('SN.ITK.DEFC.ZS'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c, date, SN.ITK.DEFC.ZS)

mmt_df <-  wb_data(country="all", 
              indicator=c('SH.STA.MMRT'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('SH.STA.MMRT'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c,  SH.STA.MMRT)


lpov_df <-  wb_data(country="all", 
              indicator=c('SE.LPV.PRIM'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('SE.LPV.PRIM'), .direction = 'downup') %>%
  filter(date==end_date)  %>%
  select(iso3c, date, SE.LPV.PRIM)  


wbl_df <- wb_data(country="all", 
              indicator=c('SG.LAW.INDX'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('SG.LAW.INDX'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c, date, SG.LAW.INDX)

safely_df <- wb_data(country="all", 
              indicator=c('SH.H2O.SMDW.ZS'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('SH.H2O.SMDW.ZS'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c, date, SH.H2O.SMDW.ZS)
  
elect_df <- wb_data(country="all", 
              indicator=c('EG.ELC.ACCS.ZS'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('EG.ELC.ACCS.ZS'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c, date, EG.ELC.ACCS.ZS)
  
manuf_df <- wb_data(country="all", 
              indicator=c('NV.IND.MANF.ZS'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('NV.IND.MANF.ZS'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c, date, NV.IND.MANF.ZS)


slums_df <- wb_data(country="all", 
              indicator=c('EN.POP.SLUM.UR.ZS'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('EN.POP.SLUM.UR.ZS'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c,  EN.POP.SLUM.UR.ZS)

ff_sub_df <- read_csv(paste0(raw_dir, "/Subsidies.csv")) %>%
  rename(iso3c=code)
  

ghg_df <- wb_data(country="all", 
              indicator=c('EN.ATM.GHGT.KT.CE'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  right_join(spi_df_empty) %>%
  fill(c('EN.ATM.GHGT.KT.CE'), .direction = 'downup') %>%
  filter(date==end_date) %>%
  select(iso3c,  EN.ATM.GHGT.KT.CE)

protected_df <- wb_data(country="all", 
              indicator=c('ER.MRN.PTMR.ZS','ER.LND.PTLD.ZS'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  fill(c('ER.MRN.PTMR.ZS','ER.LND.PTLD.ZS'), .direction = 'downup') %>%

  filter(date==end_date) %>%
  select(iso3c, date, ER.MRN.PTMR.ZS,ER.LND.PTLD.ZS)

debt_df <- wb_data(country="all", 
              indicator=c('DT.TDS.DECT.EX.ZS'),
              start_date=start_date,
              end_date=end_date,
              ) %>% # fill forward 
  group_by(country) %>%
  fill(c('DT.TDS.DECT.EX.ZS'), .direction = 'downup') %>%

  filter(date==end_date) %>%
  select(iso3c, date, DT.TDS.DECT.EX.ZS)

#bring in Sachs SDG index
#https://dashboards.sdgindex.org/explorer
# accessed on 2023-07-20

sdg_index <- read_excel(path=paste0(raw_dir, "/SDR2023-data.xlsx"),
                        sheet='Backdated SDG Index'
) %>%
  transmute(iso3c=`Country Code ISO3`,
         date=year,
         sdg_index_score=`SDG Index Score`)

#bring it all together
correlates_df_end_date <- correlates_df %>%
  left_join(ge_df) %>%
  left_join(gdp_df) %>%
  left_join(hci_df) %>%
  left_join(undernourish_df) %>%
  left_join(mmt_df) %>%
  left_join(lpov_df) %>%
  left_join(wbl_df) %>%
  left_join(safely_df) %>%
  left_join(elect_df) %>%
  left_join(manuf_df) %>%
  left_join(slums_df) %>%
  left_join(ff_sub_df) %>%
  left_join(ghg_df) %>%
  left_join(protected_df) %>%
  left_join(debt_df) %>%
  left_join(sdg_index) %>%
  filter(region!="Aggregates") %>%
  select(iso3c, c('SI.POV.DDAY', 'SI.POV.GINI','GE.EST','NY.GDP.PCAP.KD','HD.HCI.OVRL','SN.ITK.DEFC.ZS','SH.STA.MMRT','SE.LPV.PRIM','SG.LAW.INDX','SH.H2O.SMDW.ZS',"EG.ELC.ACCS.ZS",'NV.IND.MANF.ZS','EN.POP.SLUM.UR.ZS','subsidies','EN.ATM.GHGT.KT.CE','ER.MRN.PTMR.ZS','ER.LND.PTLD.ZS','DT.TDS.DECT.EX.ZS', 'sdg_index_score'))


sdgs <- c('SI.POV.DDAY','SN.ITK.DEFC.ZS','SH.STA.MMRT','SE.LPV.PRIM','SG.LAW.INDX',
          'SH.H2O.SMDW.ZS','EG.ELC.ACCS.ZS','NY.GDP.PCAP.KD','NV.IND.MANF.ZS','SI.POV.GINI',
          'EN.POP.SLUM.UR.ZS','subsidies','EN.ATM.GHGT.KT.CE','ER.MRN.PTMR.ZS','ER.LND.PTLD.ZS'
          ,'GE.EST','DT.TDS.DECT.EX.ZS', 'sdg_index_score')

sdg_names <- c('SDG 1: Extreme Poverty','SDG 2: Undernourishment','SDG 3: Maternal Mortality','SDG 4: Learning Poverty','SDG 5: Women, Business, Law Index',
          'SDG 6: Safely Managed Water','SDG 7: Access to Electricity','SDG 8: GDP per capita (2015 constant $)','SDG 9: Manufacturing value added (% of GDP)','SDG 10: Gini Index',
          'SDG 11: Population in Slums','SDG 12: Fossil Fuel Subsidies (% of GDP)','SDG 13: Greenhouse Gas Emissions','SDG 14: Marine protected areas','SDG 15: Terrestrial Protected Areas','SDG 16: Government Effectiveness','SDG 17: Total Debt Service', 'SDR: SDG Index Overall Score')
```


```{r comparison, include=FALSE}



#pull SCI values

sci_df <- read_excel(paste0(raw_dir, "Statistical_Capacity_Indicators.xlsx")) %>%
  mutate(YR2004=as.numeric(YR2004),
         iso3c=`Country Code`) %>%
  filter(`Series Code`=="IQ.SCI.OVRL") %>%
  select(-`Country Code`, -`Series Code`) %>%
  pivot_longer(
    cols=c(YR2004:YR2020),
    names_to='YR',
    values_to='SCI'
  ) %>%
  mutate(date=str_remove_all(YR,"YR"),
         date=as.numeric(date)) %>%
  select(iso3c, date, SCI)
  
  
  

# onvert it to a data frame.
sci_df <- sci_df %>%
  left_join(spi_df_empty) %>% #add on country metadata
  filter(!is.na(income)) %>%
  select(iso3c, date, SCI  ) %>%
  group_by(date) %>%
  arrange(-SCI) %>%
  mutate(SCI_rank=rank(-SCI),
         SCI_rank=if_else(is.na(SCI),as.numeric(NA),SCI_rank)) 
  
#write to csv
write_excel_csv(sci_df, paste0(output_dir, "/SCI_formatted_data.csv"))


#read in ODIN data
  for (i in c(2015,2016,2017,2018,2020,2022)) {
   temp <- read_csv(paste(raw_dir, '/','ODIN_',i,'.csv', sep=""))  %>%
        as_tibble(.name_repair = 'universal') %>%
        mutate(date=i) %>%
        filter(Data.categories=='All Categories') %>%
        select(-Year)


    assign(paste('odin_df',i,sep="_"), temp)
  }

    #bind different years together
odin_df <- bind_rows(odin_df_2015, odin_df_2016, odin_df_2017, odin_df_2018, odin_df_2020, odin_df_2022)


odin_df <- odin_df %>%
    select(Country.Code, date, Overall.score) %>%
    rename(iso3c=Country.Code,
           ODIN_score=Overall.score) %>%
    group_by(date) %>%
    arrange(-ODIN_score) %>%
    mutate(ODIN_rank=rank(-ODIN_score, na.last='NA')) %>% 
  right_join(spi_df_empty) %>%
  arrange(country, date) %>%
  group_by(country) %>%
  mutate(across(starts_with("ODIN"), na.locf, na.rm=FALSE)) %>%
  left_join(wbstats::wb_countries()) %>%
  select(iso3c, date, starts_with("ODIN"))

#write to csv
write_excel_csv(odin_df, paste0(output_dir, "/ODIN_formatted_data.csv"))

#read in the ODB
odb_df_2017 <- read_excel(path=paste0(raw_dir, '/Open Data Barometer - Historical Data (All five Editions) - Public.xlsx'),
                     sheet='ODB-2017-Rankings') %>%
  transmute(
    iso3c=ISO3,
    odb=`ODB-Score`
  )


odb_df_2016_info <- read_excel(path=paste0(raw_dir, '/Open Data Barometer - Historical Data (All five Editions) - Public.xlsx'),
                     sheet='ODB-2016-Rankings') %>%
  select(ISO3, Country) 

odb_df_2016 <- read_excel(path=paste0(raw_dir, '/Open Data Barometer - Historical Data (All five Editions) - Public.xlsx'),
                     sheet='ODB-2016-Absolute-Recalculated') %>%
  left_join(odb_df_2016_info) %>%
  transmute(
    iso3c=ISO3,
    odb_2016=`ODB Score (0-100)`
  )

odb_df_2015 <- read_excel(path=paste0(raw_dir, '/Open Data Barometer - Historical Data (All five Editions) - Public.xlsx'),
                     sheet='ODB-2015-Rankings')
odb_df_2014 <- read_excel(path=paste0(raw_dir, '/Open Data Barometer - Historical Data (All five Editions) - Public.xlsx'),
                     sheet='ODB-2014-Rankings')
odb_df_2013 <- read_excel(path=paste0(raw_dir, '/Open Data Barometer - Historical Data (All five Editions) - Public.xlsx'),
                     sheet='ODB-2013-Rankings')

odb_df <- bind_rows(
  (odb_df_2017 %>% mutate(date=2017)),
  (odb_df_2016 %>% mutate(date=2016, odb=odb_2016)),
  (odb_df_2015 %>% mutate(date=2015, iso3c=ISO3, odb=`ODB-Score-Scaled`)),
  (odb_df_2014 %>% mutate(date=2014, iso3c=ISO3, odb=`ODB-Score-Scaled`)),
  (odb_df_2013 %>% mutate(date=2013, iso3c=ISO3, odb=`ODB-Score-Scaled`)),
  ) %>%
  select(iso3c, date, odb) %>%
  left_join(country_list)

#write to csv
write_excel_csv(odb_df, paste0(output_dir, "/ODB_formatted_data.csv"))

#read in the GDB
gdb_df <- read_csv(paste0(raw_dir, "/global-data-barometer-topic--all-countries.csv")) %>%
  transmute(iso3c=iso,
            gdb=`GDB 2021 score`)

gdb_reg_df <- gdb_df %>% 
  mutate(date=2021) %>%
  select(iso3c, date, gdb) %>%
  left_join(country_list)

#write to csv
write_excel_csv(gdb_reg_df, paste0(output_dir, "/GDB_formatted_data.csv"))

#IIAG
iiag_df <- read_csv(paste0(raw_dir, "/2022 IIAG_Scores.csv")) %>%
  filter(Year==2022) %>%
  transmute(
    iso2c=Country_ISO,
    iiag_overall=`OVERALL GOVERNANCE`,
    iiag_stats=`Capacity of the Statistical System`
  ) %>%
  left_join(country_list) %>%
  select(iso3c,iiag_overall,iiag_stats )
  
#write to csv
write_excel_csv(iiag_df, paste0(output_dir, "/IIAG_formatted_data.csv"))


#create one database
comparison_df <- spi_index_df %>%
  ungroup() %>%
  filter(date ==end_date) %>%
  arrange(-SPI.INDEX) %>%
  mutate(across(starts_with('SPI.INDEX'),~1*.),
         across(starts_with('SPI.INDEX'),round,1)) %>%
  select(country, iso3c, date, income, region, SPI.INDEX) %>%
  left_join(sci_df %>% filter(date==2020) %>% ungroup() %>% select(-date)) %>%
  left_join(odin_df) %>%
  left_join(odb_df_2017) %>%
  left_join(odb_df_2016) %>%
  mutate(odb=if_else(is.na(odb),odb_2016,odb)) %>%
  left_join(gdb_df) %>%
  left_join(iiag_df) %>%
  left_join(correlates_df_end_date)

#save to output
write_excel_csv(comparison_df, paste0(output_dir, "/SPI_Index_SDG_comparisons_data.csv"))


```











```{r regdf, echo=FALSE}


predictors_df <- wbstats::wb_data(country="countries_only",
             indicator=c('NY.GDP.PCAP.KD', 'NV.IND.MANF.ZS', 'NV.AGR.TOTL.ZS','NE.TRD.GNFS.ZS', 'HD.HCI.OVRL','SE.PRM.ENRR','BN.CAB.XOKA.GD.ZS', 'CC.EST', "GE.EST", 'PV.EST', "RQ.EST", "RL.EST", "VA.EST", "BX.KLT.DINV.WD.GD.ZS", "SI.POV.DDAY", "SI.POV.GINI"),
             start_date=2010,
             end_date=2022) %>%
  mutate(date=as.numeric(date)) %>%
  left_join(sdg_index) %>%
  arrange(iso3c, date) %>%
  group_by(iso3c) %>%
  fill(c( 'NY.GDP.PCAP.KD','NV.IND.MANF.ZS', 'NV.AGR.TOTL.ZS', 'SE.PRM.ENRR', 'HD.HCI.OVRL','NE.TRD.GNFS.ZS','BN.CAB.XOKA.GD.ZS', 'CC.EST', "GE.EST", 'PV.EST', "RQ.EST", "RL.EST", "VA.EST", "BX.KLT.DINV.WD.GD.ZS","SI.POV.DDAY", "SI.POV.GINI", "sdg_index_score"), .direction="downup") %>%
  mutate(WGI.OVL=(CC.EST + GE.EST + PV.EST + RQ.EST + RL.EST + VA.EST)/6,
         lag_outcome=lag(NY.GDP.PCAP.KD),
         lag_fdi=lag(BX.KLT.DINV.WD.GD.ZS),
         lag_ge=lag(GE.EST),
         lag_poverty=lag(SI.POV.DDAY),
         lag_gini=lag(SI.POV.GINI)
         )  %>%
  ungroup()

#save to outputs
write_excel_csv(predictors_df, paste0(output_dir, "/SPI_regression_predictors.csv"))


```
















